---
name: "Ruby (Huihua) Situ"
title: "Interactions Between Collared and Not Collared White-Tailed Deer in Staten Island"
output: html_document
date: "2026-04-13"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#setwd("C:/Users/blubb/Desktop/deer")
```

Libraries
```{r}
library(tidyverse)
library(ggplot2)
library(stringr)
library(dplyr)

```


import files 
```{r}
deer <- read.csv("images_2009855.csv")
cameras <- read.csv("deployments.csv")
```




Data cleaning and prep
```{r}
deerOnly <- filter(deer, common_name == "White-tailed Deer")

#only take the useful columns to reduce the load on my computer
deerReduced <- deerOnly %>% dplyr::select(deployment_id, image_id, filename, identified_by, common_name, timestamp, number_of_objects, individual_id, behavior, highlighted, markings)

cameraReduced <- cameras %>% dplyr::select(deployment_id, subproject_name, latitude, longitude, placename, camera_name)

#remove the BLK pictures
noBLK <- deerReduced %>% filter(grepl("BLK", behavior) == FALSE)

#merge datasets together based on the camera id
deerCoord <- noBLK %>% left_join(cameraReduced, by = "deployment_id")

#add interaction column and preset to 0
deerCoord$interaction <- 0


#remove NA's
deerCoord <- na.omit(deerCoord)
```

Data fixing from older rmd file
```{r}

#creating a flag column to flag those with error for review
deerCoord$Flag <- 0

#extract the assumed correct id from the markings column and put it all in a list
ids = c()
for(x in 1:nrow(deerCoord)) {
  #take out each id using regex 
  curr <- str_extract_all(deerCoord$markings[x], "\\d+(?:-[A-Z]+)?")[[1]]
  if(length(curr) == 0) {
    ids[x] <- NA #no numbered id found
  } else {
    ids[x] <- paste(curr, collapse = ",") #paste together the ids with "," if more than one found, if only 1 this does not add a comma
  }
}

#if an issue is found it will fix it then flag it with a 1 for review
for(x in 1:nrow(deerCoord))
{ 
  #fixing the count number, it will based the count based on the annotation commas
  if (deerCoord$number_of_objects[x] != (str_count(deerCoord$markings[x], ",")+ 1))
  {
    deerCoord$number_of_objects[x] <- (str_count(deerCoord$markings[x], ",")+ 1)
    deerCoord$Flag[x] = 1
  }
  #if id column doesn't match it will update with the id from the markings column
  if(!is.na(ids[x]) && !is.na(deerCoord$individual_id[x]) &&(ids[x] !=   deerCoord$individual_id[x])) #ignoring the NAs
  {
    deerCoord$individual_id[x] <- ids[x]
    deerCoord$Flag[x] = 1
  } 
}

```


Removing bursts
```{r}
#only take one row if all the following column values are the same in multiple rows
deerDistinct <- deerCoord %>% distinct(timestamp, markings, behavior, deployment_id, subproject_name, .keep_all = TRUE)
View(deerDistinct)
```



Finding interactions of all combinations
```{r}
#finding direct interactions
for(x in 1:nrow(deerDistinct))
{
  if(grepl(",", deerDistinct$markings[x])) #only proceed if there is more than 1 deer in a picture
  {
      #split up the markings column by ,
      curr <- strsplit(deerDistinct$markings[x], ",")[[1]]
      #the regex means only take the things between the first _ and the next _
      middle <- str_match(curr,"^[^_]*_([^_]*)_.*$")[,2]
      collar <- sub("_.*", "", curr) #take anything before the first _ from curr

      #based on assumption that all collared deer are tagged
      #finding the tagged and not tagged from the curr list
      NT <- any(middle == "NT", na.rm = T) #if any item is NT this will return true
      numTag <- sum(middle != "NT" & middle != ".", na.rm = T) #return # of tagged deer
      numCol <- sum(collar == "C", na.rm = T)#return # of collared deer
      numNoCol <- sum(collar == "NC", na.rm = T)#return # of uncollared deer
      NCTag <- sum(collar == "NC" & (middle != "NT" & middle != "."), na.rm = TRUE)#of NC Tagged
      NCNT <- sum(collar == "NC" & middle == "NT", na.rm = TRUE) #return # of NC NT deer
      
      
      #if C+c that actually means 2+ collared deer 
      if (numCol > 1 && NCTag > 0 && NCNT > 0)#C + C + (NC & T) + (NC & NT)
      {
        deerDistinct$interaction[x] = 6
      }
        else if(numCol > 1 && NCTag > 0) #C + C + (NC & T)
      {
        deerDistinct$interaction[x] = 1
      } 
      else if (numCol > 1 && NCNT > 0) #C + C + (NC & NT)
      {
        deerDistinct$interaction[x] = 2
      }
      else if (numCol > 1 && numNoCol == 0) #C + C
      { 
        deerDistinct$interaction[x] = 3
      } 
      else if (numCol == 1 && NCTag > 0)#C + (NC & T)
      { 
        deerDistinct$interaction[x] = 4
      }
      else if (numCol == 1 && NCNT > 0) #C + (NC & NT)
      { 
        deerDistinct$interaction[x] = 5
      }
  }
}
```


Seperate the interactions / solo deer
```{r}
intOnly <- deerDistinct %>% filter(interaction > 0)
NoIntDeer <- deerDistinct %>% filter(interaction == 0)

```



Using time intervals to find the exact number of interactions

3 mins baseline I I I I na na na I I I 
                _______ ________ _____
                1 int     no int  2nt int
                
If no interaction for 3+ minutes after the series of photos interactions then everything before is counted as 1 interaction. 


**3 minutes**
**4 & 5 minutes** 4 and 5 gave the same result, but tiny increase of interactions from 3 mins
**1 minutes** slight increase in interactions from 3 minutes

**Test out other times by changing the 180 to a different number, all times in seconds**
```{r}
library(dplyr)

#only keep the interactions and group them 
deerOrganized <- intOnly %>% arrange(deployment_id, interaction, timestamp) %>% group_by(deployment_id, interaction)

deerInterval <- deerOrganized %>% mutate(
                prevTime = lag(timestamp), #lag = take previous
                prevMark = lag(markings), #take the previous of time and markings 
                
                #calculate the time different between the current and previous photos for all
                difTime = as.numeric(difftime(timestamp, prevTime, units = "secs")),
                #if the time is na or the marking has changed or the time is over 3 minutes then it is true its a new interaction 
                #the na is because lag assigns the first "lagged row" to NA since there is nothing before the first row 
                newInt = ifelse(is.na(difTime) | markings != prevMark | difTime > 180, 1, 0), 
                
                #makes a running total instead of just an overall total like sum()
                newEvent = cumsum(newInt)
) %>% ungroup() #ungroup since it is getting grouped differently below

View(deerInterval)

#add a new column to calculate the amount of time that specific interaction happened for
intTime3 <- deerInterval %>%
  group_by(deployment_id, newEvent, markings) %>%
  mutate(
    startTime = min(timestamp), #get the start time which will always be the smallest in each group so using min()
    endTime   = max(timestamp), 
    #as.numeric gets rid of the "secs" at the end
    interactTime = as.numeric(difftime(endTime, startTime, units = "secs"))
  )


#only keep one row from each group of interactions, 
#at this point time the original timestamp column is only really relevant for the interactions that only had a single picture (interactTime == 0) since the way it keeps rows is by pulling the first item from each group
intervalDistinct3 <- intTime3 %>% distinct(endTime, startTime, newEvent, interactTime, interaction, deployment_id, markings, .keep_all = T)

# 0 means there was a single photo of the interaction, but that does not make sense for an interaction to last 0 seconds, and the amount of zeroes can potentially cause problems in the code later on
intervalDistinct3$interactTime[intervalDistinct3$interactTime == 0] <- 1

View(intervalDistinct3)
```




Different because of time being exceeded:
```{r}
timeEx <- deerOrganized %>% mutate(
                prevTime = lag(timestamp), #lag = take previous
                prevMark = lag(markings), #take the previous of time and markings 
                
                difTime = as.numeric(difftime(timestamp, prevTime, units = "secs")),
                #if the time is na or the marking has changed or the time is over 3 minutes then it is true its a new interaction 
                newInt = ifelse(is.na(difTime) | difTime > 180, 1, 0), 
                
                newEvent = cumsum(newInt)
) %>% ungroup() #ungroup since it is getting grouped differently below

intTimetime <- timeEx %>%
  group_by(deployment_id, newEvent, markings) %>%
  mutate(
    startTime = min(timestamp), 
    endTime   = max(timestamp), 
    #as.numeric gets rid of the "secs" at the end
    interactTime = as.numeric(difftime(endTime, startTime, units = "secs"))
  )

timeExceed <- intTimetime %>% distinct(endTime, startTime, newEvent, interactTime, interaction, deployment_id, markings, .keep_all = T)
                    
View(timeExceed)

```


new interaction bcz of id changes:
```{r}
deerIntervalID <- deerOrganized %>% mutate(
                prevTime = lag(timestamp), #lag = take previous
                prevMark = lag(markings), #take the previous of time and markings 
                
                difTime = as.numeric(difftime(timestamp, prevTime, units = "secs")),
                newInt = ifelse(is.na(difTime) | markings != prevMark, 1, 0), 
                
                newEvent = cumsum(newInt)
) %>% ungroup() #ungroup since it is getting grouped differently below

intTimeID <- deerIntervalID %>%
  group_by(deployment_id, newEvent, markings) %>%
  mutate(
    startTime = min(timestamp), 
    endTime   = max(timestamp), 
    interactTime = as.numeric(difftime(endTime, startTime, units = "secs"))
  )

idChange <- intTimeID %>% distinct(endTime, startTime, newEvent, interactTime, interaction, deployment_id, markings, .keep_all = T)
                    
View(idChange)

```




Making separate datasets for each type of interaction
This will show how much the deer interact, the more pictures of them together the more they are interacting. 

```{r}

#making separate datasets for each type of interactions
CCNcT <- intervalDistinct3 %>% filter(interaction == 1)
CCNcNt <- intervalDistinct3 %>% filter(interaction == 2)
CC <- intervalDistinct3 %>% filter(interaction == 3)
CNcT <- intervalDistinct3 %>% filter(interaction == 4)
CNcNt <- intervalDistinct3 %>% filter(interaction == 5)
CCNcNtNcT <- intervalDistinct3 %>% filter(interaction == 6)

```


Number of interactions per tagged deer
```{r}
library(dplyr)
#split the ids by comma into separate rows 
deerTally <- intervalDistinct3 %>% separate_rows(individual_id, sep = ",") %>%
  group_by(individual_id) %>%#group by id so the summarize can combine it by id
  mutate(individual_id = trimws(individual_id)) %>% #get rid of any whitespace that slipped into the id section
  filter(individual_id != "") %>% #it counted all the rows without ids as a blank id and gave it its own row so this gets rid of it
  summarise(#keep a tally of how many times each id has each interaction
    int1 = sum(interaction == 1), 
    int2 = sum(interaction == 2), 
    int3 = sum(interaction == 3), 
    int4 = sum(interaction == 4), 
    int5 = sum(interaction == 5), 
    int6 = sum(interaction == 6), 
    totalInts = int1 + int2 + int3 + int4 + int5 + int6,#get the total number of interactions
    totalTime = sum(interactTime), #total length of time that deer spent
  )
View(deerTally)

```


**EDA**
percentages of each type of interaction 
```{r}

#only 2.8% of interaction images show C + C + NC_T deer interacting
nrow(CCNcT) / nrow(intervalDistinct3)

#only 10.2% of interaction images show C + C + NC_NT deer interacting
nrow(CCNcNt) / nrow(intervalDistinct3)

#only 11.4% of interaction images show 2+ collared deer interacting (no NC deer present)
nrow(CC) / nrow(intervalDistinct3)

#31.6% of interaction images show C + NC_T deer interacting
nrow(CNcT) / nrow(intervalDistinct3)

#43.8% of interaction images show c + NC_NT deer interacting
nrow(CNcNt) / nrow(intervalDistinct3)

#0.1% show C+c+ncnt+nct
nrow(CCNcNtNcT) / nrow(intervalDistinct3)
```



bar plot for number of each type of interation
```{r}
ggplot(intervalDistinct3, aes(interaction)) + 
  geom_bar(fill="hotpink") + 
  geom_text(stat = "count", aes(label = after_stat(count), vjust = -0.5)) + 
  labs(title = "Number of Interactions Per Type", x = "Interaction Type") 

```

histogram for druation of interaction

```{r}
hist(intervalDistinct3$interactTime, breaks = 50, main = "Duration of Interactions", xlab = "Duration in Seconds", ylab = "Count", col = "magenta")

```


data prep for poisson for gender vs seasons for interaction duration
```{r}
#does gender or season effect interactions more?

#update dataset with counts grouped by M/F and season(pulled from the date not the subproject_name)
      
#find the chaarcter after the last _ and put it in a list
intervalDistinct3$gender <- sapply(strsplit(intervalDistinct3$markings, ","), function(x) {
  paste(sub(".*_", "", x), collapse = ",")
}) 


#unfortunately found an issue of interactions missing the last underscore and therefore having no written gender
for(x in 1:nrow(intervalDistinct3))
{
  curr <- strsplit(intervalDistinct3$gender[x], ",")[[1]] #split the cell
  fixedLine <- sapply(curr, function(y) #if something other than LAD found in the split fix it
    {
    if(y %in% c("L", "A", "D"))
    {
      return(y)
    }
    if(grepl("OR|Y", y))
    {
      return("L")
    }
    if(grepl("WH", y))
    {
      return("A")
    }
  })
  #if a change had to be made flag it 
  if(paste(fixedLine, collapse = ",") != intervalDistinct3$gender[x])
  {
    intervalDistinct3$Flag[x] <- 1
  }
  intervalDistinct3$gender[x] <- paste(fixedLine, collapse = ",") #put the fixed listback
}



#split and find all females/males/unknown sex 
intervalDistinct3$female <- sapply(strsplit(intervalDistinct3$gender, ","), function(x)
  {
  sum(x == "L")
})
intervalDistinct3$unknown <- sapply(strsplit(intervalDistinct3$gender, ","), function(x)
  {
  sum(x == "D")
})
intervalDistinct3$male <- sapply(strsplit(intervalDistinct3$gender, ","), function(x)
  {
  sum(x == "A")
})

#extract the month
intervalDistinct3$season <- format(as.POSIXct(intervalDistinct3$timestamp), "%B")
#map the month to season
map <- c("January" = "Winter", "February" = "Winter", "March" = "Spring", "April" = "Spring", "May" = "Spring", "June" = "Summer", "July" = "Summer", "August" = "Summer", "September" = "Fall", "October" = "Fall", "November" = "Fall", "December" = "Winter")
intervalDistinct3$season <- map[intervalDistinct3$season]


View(intervalDistinct3)
```

 negative binom for gender vs seasons for interaction duration
 
Proportion of female deers has a great effect on the length of interactions between C/NC
While seasons have no significant effect.

https://stats.oarc.ucla.edu/r/dae/zinb/ <-  explains why zero inflated negative binomial wouldnt work
```{r}
library(MASS)
#dont care about interaction = 3 bcz that is C+C exclusivly
ColNcOnly <- intervalDistinct3 %>% filter(interaction != 3) %>% mutate(Fproportion = female/number_of_objects)
View(ColNcOnly)

#regular glm showed sigficantly worse AIC at over 19k
#poisDF <- glm(interactTime ~ season + Fproportion, family = poisson, data = ColNcOnly)
#summary(poisDF)


#negative binomial improved the model signifcantly meaning we have a lot of over dispuersion 

nbDF <- glm.nb(interactTime ~ season + Fproportion,
               data = ColNcOnly)
summary(nbDF)
plot(nbDF)

library(MASS)
gammaDF <- glm(interactTime ~ season + Fproportion, family = Gamma(link = log),
               data = ColNcOnly)
summary(gammaDF)
plot(gammaDF)

gammaDF2 <- glm(interactTime ~ Fproportion, family = Gamma(link = log),
               data = ColNcOnly)
summary(gammaDF2)

gammaDF3 <- glm(interactTime ~ season , family = Gamma(link = log),
               data = ColNcOnly)
summary(gammaDF3)

#negative binomial
nb2 <- glm.nb(interactTime ~ Fproportion, 
               data = ColNcOnly)
summary(nb2)

nb3 <- glm.nb(interactTime ~ season , 
               data = ColNcOnly)
summary(nb3)

#test with anova's log likleyhood
anova(gammaDF2, gammaDF, test = "Chisq")
anova(gammaDF3, gammaDF, test = "Chisq")

logLik(gammaDF)
logLik(gammaDF2)
logLik(gammaDF3)

AIC(nbDF, gammaDF, gammaDF2, gammaDF3)

library(performance)
check_overdispersion(gammaDF)#this test still shows a lot of ovedispersion despite already using negative binomial
check_overdispersion(nbDF)#this test still shows a lot of ovedispersion despite already using negative binomial
```

Taking into account the severe amount of 1 second interactions 
```{r}
#does the proportion of female deer influence the chance of an interaction lasting longer than 1 second?
ColNcOnly$longInteract <- (ColNcOnly$interactTime > 1)
ColNcOnly$longInteract <- as.numeric(ColNcOnly$longInteract)
 
longBin <- glm(longInteract ~ season + Fproportion, family = binomial, data = ColNcOnly)
summary(longBin)
check_overdispersion(longBin)


#does the proportion of female deer increase the leangth of interactions which are already over 1 second?
#only look at interactions over 1 second
longOnly <- subset(ColNcOnly, interactTime > 1)
longModel <- glm.nb(interactTime ~ season + Fproportion, data = longOnly)
summary(longModel)
plot(longModel)
check_overdispersion(longModel)
#    Null deviance: 289.22  on 242  degrees of freedom
#Residual deviance: 276.58  on 239  degrees of freedom
#AIC: 1993.7


#gamma, slightly better
gamIntOnlyDF <- subset(ColNcOnly, interactTime > 1)
gamIntOnly <- glm(interactTime ~ season + Fproportion, family = Gamma(link = log), data = longOnly)
summary(gamIntOnly)
plot(gamIntOnly)
check_overdispersion(gamIntOnly)
check_overdispersion(longModel)

#    Null deviance: 386.77  on 242  degrees of freedom
#Residual deviance: 370.04  on 239  degrees of freedom
#AIC: 1984.5
```






https://bhaskarvk.github.io/user2017.geodataviz/notebooks/03-Interactive-Maps.nb.html#using_leaflet 
https://rstudio.github.io/leaflet/articles/popups.html
https://r-graph-gallery.com/package/leaflet.html
```{r}
library(ggspatial)
library(ggmap)
library(leaflet)
library(htmltools)

#convert to dataframe and rename
ColCol <- data.frame(lng = c(CC$longitude), lat = c(CC$latitude))
COlColNcNt <- data.frame(lng = c(CCNcNt$longitude), lat = c(CCNcNt$latitude))
ColColNcT <- data.frame(lng = c(CCNcT$longitude), lat = c(CCNcT$latitude))
ColNcT <- data.frame(lng = c(CNcT$longitude), lat = c(CNcT$latitude))
ColNcNt <- data.frame(lng = c(CNcNt$longitude), lat = c(CNcNt$latitude))

#make a frequency column per location
COlColNcNt <- COlColNcNt%>% count(lng, lat, name = "freq")
ColCol <- ColCol %>% count(lng, lat, name = "freq")
ColNcNt <- ColNcNt%>% count(lng, lat, name = "freq")
ColNcT <- ColNcT%>% count(lng, lat, name = "freq")
ColColNcT <- ColColNcT %>% count(lng, lat, name = "freq")

#combine all the ones that are Collared + No collar interactions
ColNoColFull <- intervalDistinct3 %>% filter(interaction != 3)
ColNoColFull <- data.frame(lng = c(ColNoColFull$longitude), lat = c(ColNoColFull$latitude))
ColNoColFreq <- ColNoColFull %>% count(lng, lat, name = "freq")

```

Find areas with high amounts of Collared and Non-collared interaction
```{r}
pal <- colorNumeric(palette = c("lightpink", "maroon", "red"), 
                    domain = ColNoColFreq$freq)
ColNoCol_map <- leaflet() %>%
  setView(lat = 40.588176, lng = -74.141917, zoom=13.49) %>%
  addTiles(group="OSM") %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group="Dark") %>%
  addProviderTiles(providers$CartoDB.Positron, group="Light") %>%
  addLayersControl(baseGroups=c('OSM','Dark','Light')) %>%
  addCircleMarkers(data = ColNoColFreq, color = ~pal(ColNoColFreq$freq), popup = ~htmlEscape(ColNoColFreq$freq), opacity = 0.5)%>%
  addLegend("bottomright", pal = pal, values = ColNoColFreq$freq, opacity = 1)

ColNoCol_map
```


Day vs Night 
```{r}
colNoColTime <- intervalDistinct3 %>% filter(interaction != 3)
HourOfday <- colNoColTime %>% mutate(start_time = as.POSIXct(startTime), hour = as.numeric(format(start_time, "%H")), period = ifelse(hour > 6 & hour <18, "Day", "Night"))

Day <- HourOfday %>% filter(period == "Day")
Night <- HourOfday %>% filter(period == "Night")


DayDF <- data.frame(lng = c(Day$longitude), lat = c(Day$latitude))
DayFreq <- DayDF %>% count(lng, lat, name = "freq")

nightDF <- data.frame(lng = c(Night$longitude), lat = c(Night$latitude))
nightFreq <- nightDF %>% count(lng, lat, name = "freq")

############################################################################################################

#Day = red
colors <- colorNumeric(palette = c("lightpink", "maroon", "red3"), 
                    domain = DayFreq$freq)
DayFreq_map <- leaflet() %>%
  setView(lat = 40.588176, lng = -74.141917, zoom=13.49) %>%
  addTiles(group="OSM") %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group="Dark") %>%
  addProviderTiles(providers$CartoDB.Positron, group="Light") %>%
  addLayersControl(baseGroups=c('OSM','Dark','Light')) %>%
  addCircleMarkers(data = DayFreq, color = ~colors(DayFreq$freq), popup = ~htmlEscape(DayFreq$freq), opacity = 0.7)%>%
  addLegend("bottomright", pal = colors, values = DayFreq$freq, opacity = 1)

DayFreq_map
```


Night is blue 
```{r}
colors2 <- colorNumeric(palette = c("lightblue", "blue", "navy"), 
                    domain = nightFreq$freq)
nightFreq_map <- leaflet() %>%
  setView(lat = 40.588176, lng = -74.141917, zoom=13.49) %>%
  addTiles(group="OSM") %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group="Dark") %>%
  addProviderTiles(providers$CartoDB.Positron, group="Light") %>%
  addLayersControl(baseGroups=c('OSM','Dark','Light')) %>%
  addCircleMarkers(data = nightFreq, color = ~colors2(nightFreq$freq), popup = ~htmlEscape(nightFreq$freq), opacity = 0.85)%>%
  addLegend("bottomright", pal = colors2, values = nightFreq$freq, opacity = 1)

nightFreq_map


```







```{r}
library(leaflet.extras)
library(sf)
library(mapview)
library(viridis)

library(rcrimeanalysis)
```



https://cran.r-project.org/web/packages/rcrimeanalysis/vignettes/rcrimeanalysis.html
kernel density estimator KDE
```{r}
ColNoColKDE <-  ColNoColFull %>% rename(longitude = lng, latitude = lat)

kde_map(ColNoColKDE, pts = FALSE)
```


KDE by season(does not really work because the cameras are also stagnant and have repeat interactions no matter the season)
```{r}
ColNoColSeasons <- intervalDistinct3 %>% filter(interaction != 3)

Winter <- ColNoColSeasons %>% filter(season == "Winter")
Summer <- ColNoColSeasons %>% filter(season == "Summer")
Spring <- ColNoColSeasons %>% filter(season == "Spring")


WinterKDE <-  ColNoColFull %>% rename(longitude = lng, latitude = lat)
SummerKDE <-  ColNoColFull %>% rename(longitude = lng, latitude = lat)
SpringKDE <-  ColNoColFull %>% rename(longitude = lng, latitude = lat)

kde_map(WinterKDE, pts = F)
kde_map(SummerKDE, pts = F)
kde_map(SpringKDE, pts = F)
```


https://rpubs.com/nomarpicasso/1158007
Heat map, doesnt give any new information its basically just blurring the existing data points



Map for Collared + Collared
```{r}
pal1 <- colorNumeric(palette = c("turquoise3", "blue", "navy"), 
                    domain = ColCol$freq)

ColCol_map <- leaflet() %>%
  setView(lat = 40.588176, lng = -74.141917, zoom=13.25) %>%
  addTiles(group="OSM") %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group="Dark") %>%
  addProviderTiles(providers$CartoDB.Positron, group="Light") %>%
  addLayersControl(baseGroups=c('OSM','Dark','Light')) %>%
  addCircleMarkers(data = ColCol, color = ~pal1(ColCol$freq), popup = ~htmlEscape(ColCol$freq)) %>% 
  addLegend("bottomright", pal = pal1, values = ColCol$freq, opacity = 1)
ColCol_map

```


Map for Collared + Collared + No Collar&No tag
```{r}
pal2 <- colorNumeric(palette = c("turquoise3", "blue", "navy"), 
                    domain = COlColNcNt$freq)

COlColNcNt_map <- leaflet() %>%
  setView(lat = 40.588176, lng = -74.141917, zoom=13) %>%
  addTiles(group="OSM") %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group="Dark") %>%
  addProviderTiles(providers$CartoDB.Positron, group="Light") %>%
  addLayersControl(baseGroups=c('OSM','Dark','Light')) %>%
  addCircleMarkers(data = COlColNcNt, color = ~pal2(COlColNcNt$freq), popup = ~htmlEscape(COlColNcNt$freq)) %>%
  addLegend("bottomright", pal = pal2, values = COlColNcNt$freq, opacity = 1)
COlColNcNt_map

```

Map for Collared + Collared + No Collar&Tagged
```{r}
pal3 <- colorNumeric(palette = c("turquoise3", "blue", "navy"), 
                     domain = ColColNcT$freq)

ColColNcT_map <- leaflet() %>%
  setView(lat = 40.588176, lng = -74.141917, zoom=13.25) %>%
  addTiles(group="OSM") %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group="Dark") %>%
  addProviderTiles(providers$CartoDB.Positron, group="Light") %>%
  addLayersControl(baseGroups=c('OSM','Dark','Light')) %>%
  addCircleMarkers(data = ColColNcT, color = ~pal3(ColColNcT$freq), popup = ~htmlEscape(ColColNcT$freq)) %>%
  addLegend("bottomright", pal = pal3, values = ColColNcT$freq, opacity = 1)
ColColNcT_map

```


Map for Collared + No Collar&Tagged
```{r}
pal4 <- colorNumeric(palette = c("turquoise3", "blue", "navy"), 
                     domain = ColNcT$freq)

ColNcT_map <- leaflet() %>%
  setView(lat = 40.588176, lng = -74.141917, zoom=13.25) %>%
  addTiles(group="OSM") %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group="Dark") %>%
  addProviderTiles(providers$CartoDB.Positron, group="Light") %>%
  addLayersControl(baseGroups=c('OSM','Dark','Light')) %>%
  addCircleMarkers(data = ColNcT, color = ~pal4(ColNcT$freq), popup = ~htmlEscape(ColNcT$freq)) %>%
  addLegend("bottomright", pal = pal4, values = ColNcT$freq, opacity = 1)
ColNcT_map

```


Map for Collared + No Collar&No Tag
```{r}
pal5 <- colorNumeric(palette = c("turquoise3", "blue", "navy"),
                     domain = ColNcNt$freq)

ColNcNt_map <- leaflet() %>%
  setView(lat = 40.588176, lng = -74.141917, zoom=13.25) %>%
  addTiles(group="OSM") %>%
  addProviderTiles(providers$CartoDB.DarkMatter, group="Dark") %>%
  addProviderTiles(providers$CartoDB.Positron, group="Light") %>%
  addLayersControl(baseGroups=c('OSM','Dark','Light')) %>%
  addCircleMarkers(data = ColNcNt, color = ~pal5(ColNcNt$freq), popup = ~htmlEscape(ColNcNt$freq))  %>%
  addLegend("bottomright", pal = pal5, values = ColNcNt$freq, opacity = 1) 
ColNcNt_map

```

